Heatmap for all genes

Plot the heatmap of all the DE TFs from E11 to E14


```{r}
setwd("C:/Users/trant/Desktop/My-repo/My-repo/data")
e12.e11.de = read.csv("e12_e11_sig_genes.txt", header = TRUE, stringsAsFactors = FALSE)
e12.e11.de = e12.e11.de[, c(1,3)]
colnames(e12.e11.de)[1] = "Symbol"

e13.e12.de = read.csv("e13_e12_sig_genes.txt" , header = TRUE, stringsAsFactors = FALSE)
e13.e12.de = e13.e12.de[, c(1,3)]
colnames(e13.e12.de)[1] = "Symbol"
e14.e13.de = read.csv("e14_e13_sig_genes.txt", header = TRUE, stringsAsFactors = FALSE)
e14.e13.de = e14.e13.de[, c(1,3)]
colnames(e14.e13.de)[1] = "Symbol"
```

Filter for only genes with abs(logFC) >= 0.5
```{r}
e12.e11.de = e12.e11.de[abs(e12.e11.de$log2FoldChange) >= 0.5,]
e13.e12.de = e13.e12.de[abs(e13.e12.de$log2FoldChange) >= 0.5,]
e14.e13.de = e14.e13.de[abs(e14.e13.de$log2FoldChange) >= 0.5,]
```


Merge data by the "Symbol" column. Missing values are replaced by "NA"
```{r}
e12.e11_e13.e12 = merge(e12.e11.de, e13.e12.de, by = "Symbol", all = TRUE)
merged_data = merge(e12.e11_e13.e12, e14.e13.de, by = "Symbol", all = TRUE)
head(merged_data)
```
Change the column names

```{r}
colnames(merged_data) = c("Symbol", "E12/E11", "E13/E12", "E14/E13")
head(merged_data)
```

```{r}
setwd("C:/Users/trant/Desktop/My-repo/My-repo/data")
write.csv(merged_data, "all de genes.csv")
```

```{r}
merged_data_no_tf = merged_data[!(merged_data$Symbol) %in% new_order,]
head(merged_data_no_tf)
```

```{r}
setwd("C:/Users/trant/Desktop/My-repo/My-repo/data")
write.csv(merged_data_no_tf, "all de genes no tf cutoff 0.5.csv")
```

Function to convert positive logFC to 1 and negative logFC to -1. NA is assigned to 0

```{r}
up.down = function(logfc){
  if(is.na(logfc)){
    return(0)
  }else if(logfc < 0){
    return(-1)
  }else{return(1)}
}
```

Create a data frame suitable for plotting

```{r}
df.for.heatmap = data.frame(merged_data$Symbol)
df.for.heatmap$'E12/E11' = sapply(merged_data[['E12/E11']], up.down)
df.for.heatmap$'E13/E12' = sapply(merged_data[['E13/E12']], up.down)
df.for.heatmap$'E14/E13' = sapply(merged_data[['E14/E13']], up.down)
head(df.for.heatmap)
```
write this table to file
```{r}
write.csv(df.for.heatmap, "df_for_heatmap.csv")
```

Load necessary library

```{r}
library(ggplot2)
library(reshape2)
```
In order to use hclust to organize the order of gene, it is needed to make Symbol column
the row.names

```{r}
row.names(df.for.heatmap) = df.for.heatmap$merged_data.Symbol
df.for.heatmap = df.for.heatmap[, -1]
head(df.for.heatmap)
```

filter out the transcription factor
```{r}
df.for.heatmap.no.tf = df.for.heatmap[!(row.names(df.for.heatmap) %in% new_order),]
```

Calculate the order of gene using hclust
```{r}
row.order = hclust(dist(df.for.heatmap.no.tf))$order
df.for.heatmap.no.tf.new = df.for.heatmap.no.tf[row.order,]
```

Now melt the data frame
```{r}
df.for.heatmap.molten = melt(as.matrix(df.for.heatmap.no.tf.new))
names(df.for.heatmap.molten)[c(1,2)] = c("Protein", "Time")
```

Now plot the graph
```{r}
setwd("C:/Users/trant/Desktop/My-repo/My-repo/Graphs")
ggplot(df.for.heatmap.molten, aes(x = Time, y = Protein, fill = value)) +
  geom_raster() +
  scale_fill_distiller(palette = "RdBu") +
  ggtitle("Heatmap DE genes no tf")

ggsave("heatmap of DE genes no tf.png", height = 80, width = 10, limitsize = FALSE)
```


```{r}
collection = c()
for(i in 1:8805){
  c = paste(as.character(na.omit(df.for.heatmap.no.tf)[i,1:3]),collapse = "")
  collection = c(collection, c)
}
unique(collection)
```

```{r}
group1 = "100"
group2 = "110"
group3 = "111"
group4 = "101"
group5 = "-100"
group6 = "-1-10"
group7 = "-1-1-1"
group8 = "-10-1"
group9 = "010"
group10  = "011"
group11 = "0-10"
group12 = "0-1-1"
group13 = "001"
group14 = "00-1"
group15 = "1-11"
group16 = "1-10"
group17 = "10-1"
group18 = "11-1"
group19 = "-110"

group20 = "01-1"
group21 = "-101"

group22 = "-11-1"

group23 = "-111"
```

```{r}
extract = function(data, term){
  group_tf = c()
  for(i in 1:length(data[[1]])){
    if(paste(as.character(data[i, 1:3]), collapse = "") == term){
      group_tf = c(group_tf, row.names(data)[i])
    }
  }
  return(group_tf)
}
```

```{r}
group1_de = extract(df.for.heatmap.no.tf, term = group1)
group2_de = extract(df.for.heatmap.no.tf, term = group2)
group3_de = extract(df.for.heatmap.no.tf, term = group3)
group4_de = extract(df.for.heatmap.no.tf, term = group4)
group5_de = extract(df.for.heatmap.no.tf, term = group5)
group6_de = extract(df.for.heatmap.no.tf, term = group6)
group7_de = extract(df.for.heatmap.no.tf, term = group7)
group8_de = extract(df.for.heatmap.no.tf, term = group8)
group9_de = extract(df.for.heatmap.no.tf, term = group9)
group10_de = extract(df.for.heatmap.no.tf, term = group10)
group11_de = extract(df.for.heatmap.no.tf, term = group11)
group12_de = extract(df.for.heatmap.no.tf, term = group12)
group13_de = extract(df.for.heatmap.no.tf, term = group13)
group14_de = extract(df.for.heatmap.no.tf, term = group14)
group15_de = extract(df.for.heatmap.no.tf, term = group15)
group16_de = extract(df.for.heatmap.no.tf, term = group16)
group17_de = extract(df.for.heatmap.no.tf, term = group17)
group18_de = extract(df.for.heatmap.no.tf, term = group18)
group19_de = extract(df.for.heatmap.no.tf, term = group19)
group20_de = extract(df.for.heatmap.no.tf, term = group20)
group21_de = extract(df.for.heatmap.no.tf, term = group21)
group22_de = extract(df.for.heatmap.no.tf, term = group22)
group23_de = extract(df.for.heatmap.no.tf, term = group23)
```

```{r}
order = c(group23_de, group22_de, group21_de, group20_de, group19_de, group18_de, group17_de, group16_de, group15_de, group14_de, group13_de, group12_de, group11_de, group10_de, group9_de, group8_de, group7_de, group6_de, group5_de, group4_de, group3_de, group2_de, group1_de)
```

```{r}
df.for.heatmap.no.tf.new = df.for.heatmap.no.tf[order,]
```

```{r}
df.for.heatmap.molten = melt(as.matrix(df.for.heatmap.no.tf.new))
names(df.for.heatmap.molten)[c(1,2)] = c("Protein", "Time")
```

Now plot the graph
```{r}
ggplot(df.for.heatmap.molten, aes(x = Time, y = Protein, fill = value)) +
  geom_raster() +
  scale_fill_distiller(palette = "RdBu") +
  ggtitle("Heatmap of DE genes")
setwd("C:/Users/trant/Desktop/My-repo/My-repo/Graphs")
ggsave("DE heatmap arranged no tf.png", height = 80, width = 10, limitsize = FALSE)
```

```{r}
setwd("C:/Users/trant/Desktop/My-repo/My-repo/data")
all_de_genes = read.csv("all de genes no tf cutoff 0.5.csv", row.names = 2, stringsAsFactors = FALSE)
all_de_genes = all_de_genes[order,]
write.csv(all_de_genes, "all de genes no tf cut off 0.5 ordered.csv")
```


