```{r}
library(DESeq2)


```


```{r}
setwd("~/Desktop/My-repo-master/My-repo/data")
embryonic_data = read.csv("gene count matrix remove e13 sample 6.txt", row.names = 1,
                          stringsAsFactors = FALSE)
head(embryonic_data)
```

```{r}
colnames(embryonic_data) = c("e11.sample1", "e11.sample2", "e11.sample3", "e11.sample4", "e11.sample5", "e12.sample1", "e12.sample2", "e12.sample3", "e12.sample4", "e12.sample5", "e13.sample1", "e13.sample2", "e13.sample3", "e13.sample4", "e13.sample5", "e13.sample6", "e13.sample7", "e13.sample8", "e13.sample9", "e14.sample1", "e14.sample2", "e14.sample3", "e14.sample4", "e14.sample5")
```

```{r}
setwd("~/Desktop/My-repo-master/My-repo/data")
write.csv(embryonic_data, "gene count matrix remove e13 sample 6.txt")
```

```{r}
stage = c(rep("E11", 5), rep("E12", 5), rep("E13", 9), rep("E14", 5))
meta = data.frame(stage)
row.names(meta) = colnames(embryonic_data)
```

```{r}
match(colnames(embryonic_data), row.names(meta))
```


Calculate the differential expression between 2 successive time points: E12/E11, E13/E12, E14/E13
============================================================================================
Calculate differential expression of E12 vs E11
============================================================================================

First, extract the E11-E12 read counts
```{r}
e12e11 = embryonic_data[,c(1:10)]
metae12e11 = meta[c(1:10),]
```


Create DESeq object
```{r}
deseq_data_e12_e11 = DESeqDataSetFromMatrix(countData = embryonic_data[,1:10],
                                    colData = meta[1:10,],
                                    design = ~ stage)
```

Filter out genes with low counts. Only keep those with a sum of read count across all samples that are greater than 10.

```{r}
deseq_data_e12_e11 = deseq_data_e12_e11[rowSums(counts(deseq_data_e12_e11))>10, ]
```

Calculate differential expression and shrink the log fold change

```{r}
dde_e12_e11 = DESeq(deseq_data_e12_e11)
res_e12_e11 = results(dde_e12_e11, contrast = c("stage", "E12", "E11"))
res_e12_e11 = lfcShrink(dde_e12_e11, contrast = c("stage", "E12", "E11"), res = res_e12_e11)
```

extract the data and subset for differentially expressed genes

```{r}
res.all_e12_e11 = data.frame(res_e12_e11)
res.sig_e12_e11 = subset(res.all_e12_e11, padj < 0.05)
```

Write data to file
```{r}
write.csv(res.sig_e12_e11, "e12_e11_sig_genes.txt")
```

============================================================================================
Calculate differential expression between E13 vs E12
============================================================================================

```{r}
e13_e12_count_table = embryonic_data[, 6:19]
head(e13_e12_count_table)
```
Create meta data for E12 and E13 data
```{r}
stage_E12_E13 = c(rep("E12", 5), rep("E13", 9))
meta_data_E12_E13 = data.frame(stage_E12_E13)
row.names(meta_data_E12_E13) = colnames(e13_e12_count_table)
meta_data_E12_E13
```

Create DESeq object
```{r}
deseq_data_e13_e12 = DESeqDataSetFromMatrix(countData = e13_e12_count_table,
                                    colData = meta_data_E12_E13,
                                    design = ~stage_E12_E13)
```

Filter out genes with low counts. Only keep those with a sum of read count across all samples that are greater than 10.

```{r}
deseq_data_e13_e12 = deseq_data_e13_e12[rowSums(counts(deseq_data_e13_e12))>10, ]
```

Calculate differential expression and shrink the log fold change

```{r}
dde_e13_e12 = DESeq(deseq_data_e13_e12)
res_e13_e12 = results(dde_e13_e12, contrast = c("stage_E12_E13", "E13", "E12"))
res_e13_e12 = lfcShrink(dde_e13_e12, contrast = c("stage_E12_E13", "E13", "E12"), res = res_e13_e12)
```
extract the data and subset for differentially expressed genes

```{r}
res.all_e13_e12 = data.frame(res_e13_e12)
res.sig_e13_e12 = subset(res.all_e13_e12, padj < 0.05)
```

Write data to file
```{r}
write.csv(res.sig_e13_e12, "e13_e12_sig_genes.txt")
```

============================================================================================
Calculate differential expression between E14 vs E13
============================================================================================

```{r}
e14_e13_count_table = embryonic_data[, 1:10]
head(e14_e13_count_table)
```
Create meta data for E12 and E13 data
```{r}
stage_E13_E14 = c(rep("E11", 5), rep("E12", 5))
meta_data_E13_E14 = data.frame(stage_E13_E14)
row.names(meta_data_E13_E14) = colnames(e14_e13_count_table)
meta_data_E13_E14
```

Create DESeq object
```{r}
deseq_data_e14_e13 = DESeqDataSetFromMatrix(countData = e14_e13_count_table,
                                    colData = meta_data_E13_E14,
                                    design = ~stage_E13_E14)
```

Filter out genes with low counts. Only keep those with a sum of read count across all samples that are greater than 10.

```{r}
deseq_data_e14_e13 = deseq_data_e14_e13[rowSums(counts(deseq_data_e14_e13))>10, ]
```

Calculate differential expression and shrink the log fold change

```{r}
dde_e14_e13 = DESeq(deseq_data_e14_e13)
res_e14_e13 = results(dde_e14_e13, contrast = c("stage_E13_E14", "E14", "E13"))
res_e14_e13 = lfcShrink(dde_e14_e13, contrast = c("stage_E13_E14", "E14", "E13"), res = res_e14_e13)
```
extract the data and subset for differentially expressed genes

```{r}
res.all_e14_e13 = data.frame(res_e14_e13)
res.sig_e14_e13 = subset(res.all_e14_e13, padj < 0.05)
```

Write data to file
```{r}
write.csv(res.sig_e14_e13, "e14_e13_sig_genes.txt")
```


Between E12 and E11, there are 6956 DE genes
Between E13 and E12, there are 2932 DE genes
Between E14 and E13, there are 1858 DE genes
Make a table summarizing the results

```{r}
Time = c("E12 over E11", "E13 over E12", "E14 over E13")
Number_of_DE_genes = c(6956, 2932, 1858)
DE_summary = data.frame(Time, Number_of_DE_genes)
DE_summary
```

Write to file
```{r}
DE_summary_E11_E14 = write.csv(DE_summary, "DE_summary_E11_E14.txt")
```

Function to get entrezID from symbol
Parameter:
  data: dataframe with gene symbol as row.names.
  keys: vector of symbols
```{r}
symbol_entrezID = function(data, keys){
  require(org.Mm.eg.db)
  require(tidyr)
  hash.table = drop_na(AnnotationDbi::select(org.Mm.eg.db,
                                   keys = keys,
                                   keytype = "SYMBOL",
                                   columns = "ENTREZID"))
  hash.table = drop_na(hash.table)
  data_subset = data[hash.table$SYMBOL, ]
  data_subset$EntrezID = as.character(hash.table$ENTREZID)
  data_subset$Symbol = row.names(data_subset)
  row.names(data_subset) = data_subset$EntrezID
  
  return(data_subset)
}
```

Read in the list of TF
```{r}
setwd("C:/Users/trant/Desktop/My-repo/My-repo/data")
tf.db = read.csv("Mus_musculus_TF.txt", sep = "\t", stringsAsFactors = FALSE)
head(tf.db)

```

============================================================================================
Determine DE TFs between E12 and E11
============================================================================================

Convert gene symbols into EntrezID and make EntrezID the row.names
```{r}
entrezID.res.sig_e12_e11 = symbol_entrezID(data = res.sig_e12_e11, keys = row.names(res.sig_e12_e11))
```

Filter for transcription factors in all DE genes of E12/E11

```{r}
de.tf.e12.e11 = c()
for(gene in entrezID.res.sig_e12_e11$EntrezID){
  if(gene %in% tf.db$Entrez.ID){
    de.tf.e12.e11 = c(de.tf.e12.e11, gene)
  }
}
```

Subset for transcription factor only
```{r}
e12.e11.de.tf = entrezID.res.sig_e12_e11[de.tf.e12.e11,]
```

There is a total of 613 transcription factors that are differentially expressed between E12
and E11
Write this result to file
```{r}
setwd("C:/Users/trant/Desktop/My-repo/My-repo/data")
write.csv(e12.e11.de.tf, "e12_e11_all_de_tf.csv")
```

Now subset for DE TF that have logFC >= 0.5

```{r}
e12.e11.de.tf$abslogFC = abs(e12.e11.de.tf$log2FoldChange)
e12.e11.de.tf.0.5 = subset(e12.e11.de.tf, abslogFC >= 0.5)
```

There are 286 DE TFs that have log2FC either >= 0.5 or <= -0.5. Write this result to file

```{r}
setwd("C:/Users/trant/Desktop/My-repo/My-repo/data")
write.csv(e12.e11.de.tf.0.5, "e12_e11_de_tf_0.5.csv")
```

============================================================================================
Determine DE TFs between E13 and E12
============================================================================================


Convert gene symbols into EntrezID and make EntrezID the row.names
```{r}
entrezID.res.sig_e13_e12 = symbol_entrezID(data = res.sig_e13_e12, keys = row.names(res.sig_e13_e12))
```

Filter for transcription factors in all DE genes of E13/E12

```{r}
de.tf.e13.e12 = c()
for(gene in entrezID.res.sig_e13_e12$EntrezID){
  if(gene %in% tf.db$Entrez.ID){
    de.tf.e13.e12 = c(de.tf.e13.e12, gene)
  }
}
```

Subset for transcription factor only
```{r}
e13.e12.de.tf = entrezID.res.sig_e13_e12[de.tf.e13.e12,]
```

There is a total of 441 transcription factors that are differentially expressed between E13
and E12
Write this result to file
```{r}
setwd("C:/Users/trant/Desktop/My-repo/My-repo/data")
write.csv(e13.e12.de.tf, "e13_e12_all_de_tf.csv")
```

Now subset for DE TF that have logFC >= 0.5

```{r}
e13.e12.de.tf$abslogFC = abs(e13.e12.de.tf$log2FoldChange)
e13.e12.de.tf.0.5 = subset(e13.e12.de.tf, abslogFC >= 0.5)
```

There are 148 DE TFs that have log2FC either >= 0.5 or <= -0.5. Write this result to file

```{r}
setwd("C:/Users/trant/Desktop/My-repo/My-repo/data")
write.csv(e13.e12.de.tf.0.5, "e13_e12_de_tf_0.5.csv")
```

============================================================================================
Determine DE TFs between E14 and E13
============================================================================================


Convert gene symbols into EntrezID and make EntrezID the row.names
```{r}
entrezID.res.sig_e14_e13 = symbol_entrezID(data = res.sig_e14_e13, keys = row.names(res.sig_e14_e13))
```

Filter for transcription factors in all DE genes of E12/E11

```{r}
de.tf.e14.e13 = c()
for(gene in entrezID.res.sig_e14_e13$EntrezID){
  if(gene %in% tf.db$Entrez.ID){
    de.tf.e14.e13 = c(de.tf.e14.e13, gene)
  }
}
```

Subset for transcription factor only
```{r}
e14.e13.de.tf = entrezID.res.sig_e14_e13[de.tf.e14.e13,]
```

There is a total of 148 transcription factors that are differentially expressed between E14
and E13
Write this result to file
```{r}
setwd("C:/Users/trant/Desktop/My-repo/My-repo/data")
write.csv(e14.e13.de.tf, "e14_e13_all_de_tf.csv")
```

Now subset for DE TF that have logFC >= 0.5

```{r}
e14.e13.de.tf$abslogFC = abs(e14.e13.de.tf$log2FoldChange)
e14.e13.de.tf.0.5 = subset(e14.e13.de.tf, abslogFC >= 0.5)
```

There are 89 DE TFs that have log2FC either >= 0.5 or <= -0.5. Write this result to file

```{r}
setwd("C:/Users/trant/Desktop/My-repo/My-repo/data")
write.csv(e14.e13.de.tf.0.5, "e14_e13_de_tf_0.5.csv")
```

