```{r}
library(DESeq2)


```


```{r}
setwd("C:/Users/trant/Desktop/My-repo/My-repo/data")
embryonic_data = read.csv("corrected gene count matrix.txt", row.names = 1, stringsAsFactors = FALSE)
head(embryonic_data)
```

```{r}
stage = c(rep("E11",5), rep("E12",7), rep("E13", 7), rep("E14",5))
meta_data = data.frame(stage)
row.names(meta_data) = colnames(embryonic_data)
head(meta_data)
```

```{r}
dds = DESeqDataSetFromMatrix(countData = embryonic_data, colData = meta_data,
                             design = ~stage)
```

```{r}
dds_filtered = dds[rowSums(counts(dds))>10,]
```

```{r}
rld = rlog(dds_filtered, blind = FALSE)
logTransCounts = assay(rld)
```

Calculating principle component

```{r}
OGPCAN <-prcomp(logTransCounts, center = T, scale. = T)
OGPCAN_matrix <- as.data.frame(OGPCAN$rotation)
OGPCAN_matrix <- OGPCAN_matrix[, 1:2]
OGPCAN_matrix$Stage <- c(rep("E11", 5), rep("E12", 7), rep("E13", 7), rep("E14", 5))

```


```{r}
setwd("C:/Users/trant/Desktop/My-repo/My-repo/Graphs")
#Plot PCA
library(ggplot2)
ggplot(OGPCAN_matrix, aes(PC2, PC1, color = Stage)) + geom_point(size = 3) + theme(axis.text.x = element_text(size = 14, color = "black"), axis.title.x = element_text(size = 16, face = "bold"), axis.text.y = element_text(color = "black", size = 14), axis.title.y = element_text(size = 16, face = "bold"), legend.title = element_text(size = 16, face = 'bold'), legend.text = element_text(size = 14)) + scale_color_discrete(name = "Stage") + geom_label(aes(label = row.names(OGPCAN_matrix)))
ggsave("PCA plot.png", width = 20, height = 10, units = "in")
```

colnames(embryonic_data) = c("e11.sample1", "e11.sample2", "e11.sample3", "e11.sample4", "e11.sample5", "e12.sample1", "e12.sample2", "e12.sample3", "e13.sample7", "e12.sample4", "e12.sample5", "e13.sample1", "e12.sample6", "e12.sample7", "e13.sample2", "e13.sample3", "e13.sample4", "e13.sample5", "e13.sample6", "e14.sample1", "e14.sample2", "e14.sample3", "e14.sample4", "e14.sample5")



embryonic_data_2 = embryonic_data[, c(1,2,3,4,5,6,7,8,10, 11, 13, 14, 12, 15, 16, 17, 18, 19, 9, 20, 21, 22, 23, 24)]


setwd("~/Desktop/My-repo-master/My-repo/My-repo/data") 
write.csv(embryonic_data_2, "corrected gene count matrix.txt")


Calculate the differential expression between 2 successive time points: E12/E11, E13/E12, E14/E13
============================================================================================
Calculate differential expression of E12 vs E11
============================================================================================

First, extract the E11-E12 read counts
```{r}
e12e11 = embryonic_data[,c(1:12)]
stage= c(rep("E11", 5), rep("E12", 7))
meta_e12_e11 = data.frame(stage)
row.names(meta_e12_e11) = colnames(e12e11)
```


Create DESeq object
```{r}
deseq_data_e12_e11 = DESeqDataSetFromMatrix(countData = e12e11,
                                    colData = meta_e12_e11,
                                    design = ~ stage)
```

Filter out genes with low counts. Only keep those with a sum of read count across all samples that are greater than 10.

```{r}
deseq_data_e12_e11 = deseq_data_e12_e11[rowSums(counts(deseq_data_e12_e11))>10, ]
```

Calculate differential expression and shrink the log fold change

```{r}
dde_e12_e11 = DESeq(deseq_data_e12_e11)
res_e12_e11 = results(dde_e12_e11, contrast = c("stage", "E12", "E11"))
res_e12_e11 = lfcShrink(dde_e12_e11, contrast = c("stage", "E12", "E11"), res = res_e12_e11)
```

extract the data and subset for differentially expressed genes

```{r}
res.all_e12_e11 = data.frame(res_e12_e11)
res.sig_e12_e11 = subset(res.all_e12_e11, padj < 0.05)
```

Write data to file
```{r}
setwd("C:/Users/trant/Desktop/My-repo/My-repo/data")
write.csv(res.sig_e12_e11, "e12_e11_sig_genes.txt")
```

============================================================================================
Calculate differential expression between E13 vs E12
============================================================================================

```{r}
e13_e12_count_table = embryonic_data[, 6:19]
head(e13_e12_count_table)
```
Create meta data for E12 and E13 data
```{r}
stage_E12_E13 = c(rep("E12", 7), rep("E13", 7))
meta_data_E12_E13 = data.frame(stage_E12_E13)
row.names(meta_data_E12_E13) = colnames(e13_e12_count_table)
meta_data_E12_E13
```

Create DESeq object
```{r}
deseq_data_e13_e12 = DESeqDataSetFromMatrix(countData = e13_e12_count_table,
                                    colData = meta_data_E12_E13,
                                    design = ~stage_E12_E13)
```

Filter out genes with low counts. Only keep those with a sum of read count across all samples that are greater than 10.

```{r}
deseq_data_e13_e12 = deseq_data_e13_e12[rowSums(counts(deseq_data_e13_e12))>10, ]
```

Calculate differential expression and shrink the log fold change

```{r}
dde_e13_e12 = DESeq(deseq_data_e13_e12)
res_e13_e12 = results(dde_e13_e12, contrast = c("stage_E12_E13", "E13", "E12"))
res_e13_e12 = lfcShrink(dde_e13_e12, contrast = c("stage_E12_E13", "E13", "E12"), res = res_e13_e12)
```
extract the data and subset for differentially expressed genes

```{r}
res.all_e13_e12 = data.frame(res_e13_e12)
res.sig_e13_e12 = subset(res.all_e13_e12, padj < 0.05)
```

Write data to file
```{r}
setwd("C:/Users/trant/Desktop/My-repo/My-repo/data")
write.csv(res.sig_e13_e12, "e13_e12_sig_genes.txt")
```

============================================================================================
Calculate differential expression between E14 vs E13
============================================================================================

```{r}
e14_e13_count_table = embryonic_data[, 13:24]
head(e14_e13_count_table)
```
Create meta data for E12 and E13 data
```{r}
stage_E13_E14 = c(rep("E13", 7), rep("E14", 5))
meta_data_E13_E14 = data.frame(stage_E13_E14)
row.names(meta_data_E13_E14) = colnames(e14_e13_count_table)
meta_data_E13_E14
```

Create DESeq object
```{r}
deseq_data_e14_e13 = DESeqDataSetFromMatrix(countData = e14_e13_count_table,
                                    colData = meta_data_E13_E14,
                                    design = ~stage_E13_E14)
```

Filter out genes with low counts. Only keep those with a sum of read count across all samples that are greater than 10.

```{r}
deseq_data_e14_e13 = deseq_data_e14_e13[rowSums(counts(deseq_data_e14_e13))>10, ]
```

Calculate differential expression and shrink the log fold change

```{r}
dde_e14_e13 = DESeq(deseq_data_e14_e13)
res_e14_e13 = results(dde_e14_e13, contrast = c("stage_E13_E14", "E14", "E13"))
res_e14_e13 = lfcShrink(dde_e14_e13, contrast = c("stage_E13_E14", "E14", "E13"), res = res_e14_e13)
```
extract the data and subset for differentially expressed genes

```{r}
res.all_e14_e13 = data.frame(res_e14_e13)
res.sig_e14_e13 = subset(res.all_e14_e13, padj < 0.05)
```

Write data to file
```{r}
setwd("C:/Users/trant/Desktop/My-repo/My-repo/data")
write.csv(res.sig_e14_e13, "e14_e13_sig_genes.txt")
```

============================================================================================
Calculate differential expression between E13 vs E11
============================================================================================

```{r}
e13_e11_count_table = embryonic_data[, c(1:5, 13:19)]
head(e13_e11_count_table)
```
Create meta data for E12 and E13 data
```{r}
stage_E13_E11 = c(rep("E11", 5), rep("E13", 7))
meta_data_E13_E11 = data.frame(stage_E13_E11)
row.names(meta_data_E13_E11) = colnames(e13_e11_count_table)
meta_data_E13_E11
```

Create DESeq object
```{r}
deseq_data_e13_e11 = DESeqDataSetFromMatrix(countData = e13_e11_count_table,
                                    colData = meta_data_E13_E11,
                                    design = ~stage_E13_E11)
```

Filter out genes with low counts. Only keep those with a sum of read count across all samples that are greater than 10.

```{r}
deseq_data_e13_e11 = deseq_data_e13_e11[rowSums(counts(deseq_data_e13_e11))>10, ]
```

Calculate differential expression and shrink the log fold change

```{r}
dde_e13_e11 = DESeq(deseq_data_e13_e11)
res_e13_e11 = results(dde_e13_e11, contrast = c("stage_E13_E11", "E13", "E11"))
res_e13_e11 = lfcShrink(dde_e13_e11, contrast = c("stage_E13_E11", "E13", "E11"), res = res_e13_e11)
```
extract the data and subset for differentially expressed genes

```{r}
res.all_e13_e11 = data.frame(res_e13_e11)
res.sig_e13_e11 = subset(res.all_e13_e11, padj < 0.05)
```

Write data to file
```{r}
setwd("C:/Users/trant/Desktop/My-repo/My-repo/data")
write.csv(res.sig_e14_e13, "e13_e11_sig_genes.txt")
```

============================================================================================
Calculate differential expression between E14 vs E11
============================================================================================

```{r}
e14_e11_count_table = embryonic_data[, c(1:5, 20:24)]
head(e14_e11_count_table)
```
Create meta data for E12 and E13 data
```{r}
stage_E14_E11 = c(rep("E11", 5), rep("E14", 5))
meta_data_E14_E11 = data.frame(stage_E14_E11)
row.names(meta_data_E14_E11) = colnames(e14_e11_count_table)
meta_data_E14_E11
```

Create DESeq object
```{r}
deseq_data_e14_e11 = DESeqDataSetFromMatrix(countData = e14_e11_count_table,
                                    colData = meta_data_E14_E11,
                                    design = ~stage_E14_E11)
```

Filter out genes with low counts. Only keep those with a sum of read count across all samples that are greater than 10.

```{r}
deseq_data_e14_e11 = deseq_data_e14_e11[rowSums(counts(deseq_data_e14_e11))>10, ]
```

Calculate differential expression and shrink the log fold change

```{r}
dde_e14_e11 = DESeq(deseq_data_e14_e11)
res_e14_e11 = results(dde_e14_e11, contrast = c("stage_E14_E11", "E14", "E11"))
res_e14_e11 = lfcShrink(dde_e14_e11, contrast = c("stage_E14_E11", "E14", "E11"), res = res_e14_e11)
```
extract the data and subset for differentially expressed genes

```{r}
res.all_e14_e11 = data.frame(res_e14_e11)
res.sig_e14_e11 = subset(res.all_e14_e11, padj < 0.05)
```

Write data to file
```{r}
setwd("C:/Users/trant/Desktop/My-repo/My-repo/data")
write.csv(res.sig_e14_e11, "e14_e11_sig_genes.txt")
```

============================================================================================
Calculate differential expression between E14 vs E12
============================================================================================

```{r}
e14_e12_count_table = embryonic_data[, c(6:12, 20:24)]
head(e14_e12_count_table)
```
Create meta data for E12 and E13 data
```{r}
stage_E14_E12 = c(rep("E12", 7), rep("E14", 5))
meta_data_E14_E12 = data.frame(stage_E14_E12)
row.names(meta_data_E14_E12) = colnames(e14_e12_count_table)
meta_data_E14_E12
```

Create DESeq object
```{r}
deseq_data_e14_e12 = DESeqDataSetFromMatrix(countData = e14_e12_count_table,
                                    colData = meta_data_E14_E12,
                                    design = ~stage_E14_E12)
```

Filter out genes with low counts. Only keep those with a sum of read count across all samples that are greater than 10.

```{r}
deseq_data_e14_e12 = deseq_data_e14_e12[rowSums(counts(deseq_data_e14_e12))>10, ]
```

Calculate differential expression and shrink the log fold change

```{r}
dde_e14_e12 = DESeq(deseq_data_e14_e12)
res_e14_e12 = results(dde_e14_e12, contrast = c("stage_E14_E12", "E14", "E12"))
res_e14_e12 = lfcShrink(dde_e14_e12, contrast = c("stage_E14_E12", "E14", "E12"), res = res_e14_e12)
```
extract the data and subset for differentially expressed genes

```{r}
res.all_e14_e12 = data.frame(res_e14_e12)
res.sig_e14_e12 = subset(res.all_e14_e12, padj < 0.05)
```

Write data to file
```{r}
setwd("C:/Users/trant/Desktop/My-repo/My-repo/data")
write.csv(res.sig_e14_e12, "e14_e12_sig_genes.txt")
```













Between E12 and E11, there are 7762 DE genes
Between E13 and E12, there are 4458 DE genes
Between E14 and E13, there are 855 DE genes
Between E13 and E11, there are 11042 DE genes
Between E14 and E11, there are 10683 DE genes
Between E14 and E12, there are 6862 DE genes
Make a table summarizing the results

```{r}
Time = c("E12 over E11", "E13 over E12", "E14 over E13", "E13 over E11", "E14 over E11", "E14 over E12")
Number_of_DE_genes = c(7762, 4458, 855, 11042, 10683, 6862)
DE_summary = data.frame(Time, Number_of_DE_genes)
DE_summary
```

Write to file
```{r}
setwd("C:/Users/trant/Desktop/My-repo/My-repo/data")
DE_summary_E11_E14 = write.csv(DE_summary, "DE_summary_E11_E14.txt")
```
