

```{r}
setwd("C:/Users/trant/Desktop/Embryonic RNA seq/My-repo/data")
TF_cofactor = read.csv("Mus_musculus_TF_cofactors.txt", sep = "\t", header = TRUE, 
                       stringsAsFactors = FALSE)
head(TF_cofactor)
```

```{r}
e12e11 = read.csv("e12_e11_sig_genes.txt", sep = ",", header = TRUE, 
                  stringsAsFactors = FALSE, row.names = 1)
e13e12 = read.csv("e13_e12_sig_genes.txt", sep = ",", header = TRUE, 
                  stringsAsFactors = FALSE, row.names = 1)
e14e13 = read.csv("e14_e13_sig_genes.txt", sep = ",", header = TRUE, 
                  stringsAsFactors = FALSE, row.names = 1)
```


Function to get entrezID from symbol
Parameter:
  data: dataframe with gene symbol as row.names.
  keys: vector of symbols
```{r}
symbol_entrezID = function(data, keys){
  require(org.Mm.eg.db)
  require(tidyr)
  hash.table = drop_na(AnnotationDbi::select(org.Mm.eg.db,
                                   keys = keys,
                                   keytype = "SYMBOL",
                                   columns = "ENTREZID"))
  hash.table = drop_na(hash.table)
  data_subset = data[hash.table$SYMBOL, ]
  data_subset$EntrezID = as.character(hash.table$ENTREZID)
  data_subset$Symbol = row.names(data_subset)
  row.names(data_subset) = data_subset$EntrezID
  
  return(data_subset)
}
```

Convert gene symbols into EntrezID and make EntrezID the row.names
```{r}
entrezID.res.sig_e12_e11 = symbol_entrezID(data = e12e11, keys = row.names(e12e11))
head(entrezID.res.sig_e12_e11)
```
Filter for transcription cofactors in all DE genes of E12/E11

```{r}
de.cotf.e12.e11 = c()
for(gene in entrezID.res.sig_e12_e11$EntrezID){
  if(gene %in% TF_cofactor$Entrez.ID){
    de.cotf.e12.e11 = c(de.cotf.e12.e11, gene)
  }
}
```

Subset for transcription factor only
```{r}
e12.e11.de.cotf = entrezID.res.sig_e12_e11[de.cotf.e12.e11,]
```

There are 488 DE cofactors. Write this result to file
```{r}
write.csv(e12.e11.de.cotf, "e12.e11.all.de.cofactors.csv")
```

Now filter for those with log2FC >= 0.5
```{r}
e12.e11.de.cotf.0.5 = subset(e12.e11.de.cotf, abs(e12.e11.de.cotf$log2FoldChange) >= 0.25)
```

Write to file
```{r}
write.csv(e12.e11.de.cotf.0.5, "e12.e11.de.cotf.0.5.csv")
```


```{r}
extract_genes = function(de_data, reference, save1 = FALSE, file1, save2 = FALSE, file2 ){
  #Convert gene symbols into EntrezID and make EntrezID the row.names
  entrezID.res.sig = symbol_entrezID(data = de_data, keys = row.names(de_data))
  #Filter for desired genes
  de.cotf = c()
  for(gene in entrezID.res.sig$EntrezID){
    if(gene %in% reference$Entrez.ID){
      de.cotf = c(de.cotf, gene)
      }
  }
  #subset for desired genes
  de.cotf_df = entrezID.res.sig[de.cotf,]
  if(save1){
    write.csv(de.cotf_df, file1)
  }
  #subset for log2FC >= 0.5
  de.cotf.0.5 = subset(de.cotf_df, abs(de.cotf_df$log2FoldChange) >= 0.5)
  if(save2){
    write.csv(de.cotf.0.5, file2)
  }
  return(de.cotf.0.5)
}
```

```{r}
e13.e12.de.cotf.0.5 = extract_genes(de_data = e13e12, reference = TF_cofactor,
                                    save1 = TRUE, file1 = "e13.e12.all.de.cofactors.csv",
                                    save2 = TRUE, file2 = "e13.e12.de.cotf.0.5.csv")
e13.e12.de.cotf.0.5
```
```{r}
e14.e13.de.cotf.0.5 = extract_genes(de_data = e14e13, reference = TF_cofactor,
                                    save1 = TRUE, file1 = "e14.e13.all.de.cofactors.csv",
                                    save2 = TRUE, file2 = "e14.e13.de.cotf.0.5.csv")
e14.e13.de.cotf.0.5
```


