
Function to get entrezID from symbol
Parameter:
  data: dataframe with gene symbol as row.names.
  keys: vector of symbols
```{r}
symbol_entrezID = function(data, keys){
  require(org.Mm.eg.db)
  require(tidyr)
  hash.table = drop_na(AnnotationDbi::select(org.Mm.eg.db,
                                   keys = keys,
                                   keytype = "SYMBOL",
                                   columns = "ENTREZID"))
  hash.table = drop_na(hash.table)
  data_subset = data[hash.table$SYMBOL, ]
  data_subset$EntrezID = as.character(hash.table$ENTREZID)
  data_subset$Symbol = row.names(data_subset)
  row.names(data_subset) = data_subset$EntrezID
  
  return(data_subset)
}
```

Read in the list of TF
```{r}
setwd("~/Desktop/My-repo-master/My-repo/My-repo/data")
tf.db = read.csv("Mus_musculus_TF.txt", sep = "\t", stringsAsFactors = FALSE)
head(tf.db)

```

============================================================================================
Determine DE TFs between E12 and E11
============================================================================================
Read in the data frame

```{r}
setwd("~/Desktop/My-repo-master/My-repo/My-repo/data")
res.sig_e12_e11 = read.csv("e12_e11_sig_genes.txt", row.names = 1, stringsAsFactors = FALSE)
head(res.sig_e12_e11)

```

Convert gene symbols into EntrezID and make EntrezID the row.names
```{r}
entrezID.res.sig_e12_e11 = symbol_entrezID(data = res.sig_e12_e11, keys = row.names(res.sig_e12_e11))
```

Filter for transcription factors in all DE genes of E12/E11

```{r}
de.tf.e12.e11 = c()
for(gene in entrezID.res.sig_e12_e11$EntrezID){
  if(gene %in% tf.db$Entrez.ID){
    de.tf.e12.e11 = c(de.tf.e12.e11, gene)
  }
}
```

Subset for transcription factor only
```{r}
e12.e11.de.tf = entrezID.res.sig_e12_e11[de.tf.e12.e11,]
```

There is a total of 646 transcription factors that are differentially expressed between E12 and E11
Write this result to file
```{r}
setwd("~/Desktop/My-repo-master/My-repo/My-repo/data")
write.csv(e12.e11.de.tf, "e12_e11_all_de_tf.csv")
```

Now subset for DE TF that have logFC >= 1

```{r}
e12.e11.de.tf.1 = subset(e12.e11.de.tf, abs(e12.e11.de.tf$log2FoldChange) >= 1)
```

There are 140 DE TFs that have log2FC either >= 1 or <= -1. Write this result to file

```{r}
setwd("~/Desktop/My-repo-master/My-repo/My-repo/data")
write.csv(e12.e11.de.tf.1, "e12_e11_de_tf_1.csv")
```

============================================================================================
Determine DE TFs between E13 and E12
============================================================================================

```{r}
setwd("~/Desktop/My-repo-master/My-repo/My-repo/data")
res.sig_e13_e12 = read.csv("e13_e12_sig_genes.txt", row.names = 1, stringsAsFactors = FALSE)
```

Convert gene symbols into EntrezID and make EntrezID the row.names
```{r}
entrezID.res.sig_e13_e12 = symbol_entrezID(data = res.sig_e13_e12, keys = row.names(res.sig_e13_e12))
```

Filter for transcription factors in all DE genes of E13/E12

```{r}
de.tf.e13.e12 = c()
for(gene in entrezID.res.sig_e13_e12$EntrezID){
  if(gene %in% tf.db$Entrez.ID){
    de.tf.e13.e12 = c(de.tf.e13.e12, gene)
  }
}
```

Subset for transcription factor only
```{r}
e13.e12.de.tf = entrezID.res.sig_e13_e12[de.tf.e13.e12,]
```

There is a total of 437 transcription factors that are differentially expressed between E13 and E12
Write this result to file
```{r}
setwd("~/Desktop/My-repo-master/My-repo/My-repo/data")
write.csv(e13.e12.de.tf, "e13_e12_all_de_tf.csv")
```

Now subset for DE TF that have logFC >= 1

```{r}

e13.e12.de.tf.1 = subset(e13.e12.de.tf, abs(e13.e12.de.tf$log2FoldChange) >= 1)
```

There are 131 DE TFs that have log2FC either >= 0.5 or <= -0.5. Write this result to file

```{r}
setwd("~/Desktop/My-repo-master/My-repo/My-repo/data")
write.csv(e13.e12.de.tf.1, "e13_e12_de_tf_1.csv")
```

============================================================================================
Determine DE TFs between E14 and E13
============================================================================================

```{r}
setwd("~/Desktop/My-repo-master/My-repo/My-repo/data")
res.sig_e14_e13 = read.csv("e14_e13_sig_genes.txt", row.names = 1, stringsAsFactors = FALSE)
```


Convert gene symbols into EntrezID and make EntrezID the row.names
```{r}
entrezID.res.sig_e14_e13 = symbol_entrezID(data = res.sig_e14_e13, keys = row.names(res.sig_e14_e13))
```

Filter for transcription factors in all DE genes of E14/E13

```{r}
de.tf.e14.e13 = c()
for(gene in entrezID.res.sig_e14_e13$EntrezID){
  if(gene %in% tf.db$Entrez.ID){
    de.tf.e14.e13 = c(de.tf.e14.e13, gene)
  }
}
```

Subset for transcription factor only
```{r}
e14.e13.de.tf = entrezID.res.sig_e14_e13[de.tf.e14.e13,]
```

There is a total of 67 transcription factors that are differentially expressed between E14
and E13
Write this result to file
```{r}
setwd("~/Desktop/My-repo-master/My-repo/My-repo/data")
write.csv(e14.e13.de.tf, "e14_e13_all_de_tf.csv")
```

Now subset for DE TF that have logFC >= 1

```{r}
e14.e13.de.tf.1 = subset(e14.e13.de.tf, abs(e14.e13.de.tf$log2FoldChange) >= 1)
```

There are 35 DE TFs that have log2FC either >= 0.5 or <= -0.5. Write this result to file

```{r}
setwd("~/Desktop/My-repo-master/My-repo/My-repo/data")
write.csv(e14.e13.de.tf.1, "e14_e13_de_tf_1.csv")
```

============================================================================================
Determine DE TFs between E13 and E11
============================================================================================
Read in the data frame

```{r}
setwd("~/Desktop/My-repo-master/My-repo/My-repo/data")
res.sig_e13_e11 = read.csv("e13_e11_sig_genes.txt", row.names = 1, stringsAsFactors = FALSE)
head(res.sig_e13_e11)

```

Convert gene symbols into EntrezID and make EntrezID the row.names
```{r}
entrezID.res.sig_e13_e11 = symbol_entrezID(data = res.sig_e13_e11, keys = row.names(res.sig_e13_e11))
```

Filter for transcription factors in all DE genes of E13/E11

```{r}
de.tf.e13.e11 = c()
for(gene in entrezID.res.sig_e13_e11$EntrezID){
  if(gene %in% tf.db$Entrez.ID){
    de.tf.e13.e11 = c(de.tf.e13.e11, gene)
  }
}
```

Subset for transcription factor only
```{r}
e13.e11.de.tf = entrezID.res.sig_e13_e11[de.tf.e13.e11,]
```

There is a total of 67 transcription factors that are differentially expressed between E13 and E11
Write this result to file
```{r}
setwd("~/Desktop/My-repo-master/My-repo/My-repo/data")
write.csv(e13.e11.de.tf, "e13_e11_all_de_tf.csv")
```

Now subset for DE TF that have logFC >= 1

```{r}
e13.e11.de.tf.1 = subset(e13.e11.de.tf, abs(e13.e11.de.tf$log2FoldChange) >= 1)
```

There are 35 DE TFs that have log2FC either >= 0.5 or <= -0.5. Write this result to file

```{r}
setwd("~/Desktop/My-repo-master/My-repo/My-repo/data")
write.csv(e13.e11.de.tf.1, "e13_e11_de_tf_1.csv")
```

============================================================================================
Determine DE TFs between E14 and E11
============================================================================================
Read in the data frame

```{r}
setwd("~/Desktop/My-repo-master/My-repo/My-repo/data")
res.sig_e14_e11 = read.csv("e14_e11_sig_genes.txt", row.names = 1, stringsAsFactors = FALSE)
head(res.sig_e14_e11)

```

Convert gene symbols into EntrezID and make EntrezID the row.names
```{r}
entrezID.res.sig_e14_e11 = symbol_entrezID(data = res.sig_e14_e11, keys = row.names(res.sig_e14_e11))
```

Filter for transcription factors in all DE genes of E12/E11

```{r}
de.tf.e14.e11 = c()
for(gene in entrezID.res.sig_e14_e11$EntrezID){
  if(gene %in% tf.db$Entrez.ID){
    de.tf.e14.e11 = c(de.tf.e14.e11, gene)
  }
}
```

Subset for transcription factor only
```{r}
e14.e11.de.tf = entrezID.res.sig_e14_e11[de.tf.e14.e11,]
```

There is a total of 868 transcription factors that are differentially expressed between E14 and E11
Write this result to file
```{r}
setwd("~/Desktop/My-repo-master/My-repo/My-repo/data")
write.csv(e14.e11.de.tf, "e14_e11_all_de_tf.csv")
```

Now subset for DE TF that have logFC >= 1

```{r}
e14.e11.de.tf.1 = subset(e14.e11.de.tf, abs(e14.e11.de.tf$log2FoldChange) >= 1)
```

There are 653 DE TFs that have log2FC either >= 0.5 or <= -0.5. Write this result to file

```{r}
setwd("~/Desktop/My-repo-master/My-repo/My-repo/data")
write.csv(e14.e11.de.tf.1, "e14_e11_de_tf_1.csv")
```

============================================================================================
Determine DE TFs between E14 and E12
============================================================================================
Read in the data frame

```{r}
setwd("~/Desktop/My-repo-master/My-repo/My-repo/data")
res.sig_e14_e12 = read.csv("e14_e12_sig_genes.txt", row.names = 1, stringsAsFactors = FALSE)
head(res.sig_e14_e12)

```

Convert gene symbols into EntrezID and make EntrezID the row.names
```{r}
entrezID.res.sig_e14_e12 = symbol_entrezID(data = res.sig_e14_e12, keys = row.names(res.sig_e14_e12))
```

Filter for transcription factors in all DE genes of E12/E11

```{r}
de.tf.e14.e12 = c()
for(gene in entrezID.res.sig_e14_e12$EntrezID){
  if(gene %in% tf.db$Entrez.ID){
    de.tf.e14.e12 = c(de.tf.e14.e12, gene)
  }
}
```

Subset for transcription factor only
```{r}
e14.e12.de.tf = entrezID.res.sig_e14_e12[de.tf.e14.e12,]
```

There is a total of 601 transcription factors that are differentially expressed between E12 and E11
Write this result to file
```{r}
setwd("~/Desktop/My-repo-master/My-repo/My-repo/data")
write.csv(e14.e12.de.tf, "e14_e12_all_de_tf.csv")
```

Now subset for DE TF that have logFC >= 0.5

```{r}
e14.e12.de.tf.1 = subset(e14.e12.de.tf, abs(e14.e12.de.tf$log2FoldChange) >= 1)
```

There are 386 DE TFs that have log2FC either >= 0.5 or <= -0.5. Write this result to file

```{r}
setwd("~/Desktop/My-repo-master/My-repo/My-repo/data")
write.csv(e14.e12.de.tf.1, "e14_e12_de_tf_1.csv")
```
