```{r}
setwd("~/Desktop/My-repo-master/My-repo/data")
data = read.csv("all.tf.with.family.csv",
                header = TRUE,
                sep = ",",
                stringsAsFactors = FALSE,
                row.names = 1)
head(data)
```


```{r}
for(i in 1:length(data$Family)){
  if(data$Family[i] == "Homeobox" | data$Family[i] == "Pou" | data$Family[i] == "PAX"){
    data$Family[i] = "Homeobox"
  }
}

```

```{r}
for(i in 1:length(data$Family)){
  if(data$Family[i] == "Miscellaneous" | data$Family[i] == "Others"){
    data$Family[i] = "Miscellaneous"
  }
}
```


```{r}
#Group family with low counts and similar downregulation patterns together
for(i in 1:length(data$Family)){
  if(data$Family[i] == "AF-4" | data$Family[i] == "CSD" | data$Family[i] == "HMGA" | data$Family[i] == "RFX"| data$Family[i] == "Tub" | data$Family[i] == "NDT80/PhoG" | data$Family[i] == "MBD" | data$Family[i] == "E2F" | data$Family[i] == "CG1" | data$Family[i] == "CP2" | data$Family[i] == "DACH" | data$Family[i] == "MH1" | data$Family[i] == "TF_Otx" | data$Family[i] == "CG-1"){
    data$Family[i] = "downregulated, low quantity"
  }
}

```


```{r}
#Group family with low counts and similar upregulation patterns together
for(i in 1:length(data$Family)){
  if(data$Family[i] == "COE" | data$Family[i] == "CSRNP_N" | data$Family[i] == "CTF/NFI" | data$Family[i] == "ESR-like"| data$Family[i] == "HPD" | data$Family[i] == "NGFIB-like" | data$Family[i] == "Runt" | data$Family[i] == "RHD" | data$Family[i] == "SRF" | data$Family[i] == "TEA"){
    data$Family[i] = "upregulated, low quantity"
  }
}

```

```{r}
for(i in 1:length(data$Family)){
  if(data$Family[i] == "zf-C2H2" | data$Family[i] == "zf-C2HC" | data$Family[i] == "zf-CCCH" | data$Family[i] == "zf-GATA"){
    data$Family[i] = "Zinc fingers"
  }
}
```

```{r}
data_sorted = data[order(data$Family),]
```

```{r}
bHLH = subset(data_sorted, Family == "bHLH")
AP2 = subset(data_sorted, Family == "AP-2")
ARID = subset(data_sorted, Family == "ARID")
ETS = subset(data_sorted, Family == "ETS")
Fork_head = subset(data_sorted, Family == "Fork_head")
HMG = subset(data_sorted, Family == "HMG")
Homeobox = subset(data_sorted, Family == "Homeobox")
IRF = subset(data_sorted, Family == "IRF")
Tbox = subset(data_sorted, Family == "T-box")
TF_bZIP = subset(data_sorted, Family == "TF_bZIP")
THR_like = subset(data_sorted, Family == "THR-like")
ZBTB = subset(data_sorted, Family == "ZBTB")
zincfinger = subset(data_sorted, Family == "Zinc fingers")
```

```{r}
bHLH = bHLH[hclust(dist(bHLH[,1:3]))$order,]
AP2 = AP2[hclust(dist(AP2[,1:3]))$order,]
ARID = ARID[hclust(dist(ARID[,1:3]))$order,]
ETS = ETS[hclust(dist(ETS[,1:3]))$order,]
Fork_head = Fork_head[hclust(dist(Fork_head[,1:3]))$order,]
HMG = HMG[hclust(dist(HMG[,1:3]))$order,]
Homeobox = Homeobox[hclust(dist(Homeobox[,1:3]))$order,]
IRF = IRF[hclust(dist(IRF[,1:3]))$order,]
Tbox = Tbox[hclust(dist(Tbox[,1:3]))$order,]
TF_bZIP = TF_bZIP[hclust(dist(TF_bZIP[,1:3]))$order,]
THR_like = THR_like[hclust(dist(THR_like[,1:3]))$order,]
ZBTB = ZBTB[hclust(dist(ZBTB[,1:3]))$order,]
zincfinger = zincfinger[hclust(dist(zincfinger[,1:3]))$order,]
```

```{r}
rowbind = rbind(bHLH, ETS, Fork_head, Homeobox, Tbox, TF_bZIP, ZBTB, zincfinger)
```

```{r}
plotdata = data.frame(rowbind$E12.E11, rowbind$E13.E12, rowbind$E14.E13)
row.names(plotdata) = rowbind$Symbol
colnames(plotdata) = c("E12.E11", "E13.E12", "E14.E13")
```

```{r}
setwd("~/Desktop/My-repo-master/My-repo/data")
write.csv(plotdata, "clustered.csv")

```



```{r}
library(reshape2)
plotdata_melt = melt(as.matrix(plotdata))
names(plotdata_melt)[c(1,2)] = c("Protein", "Time")
```

```{r}
library(ggplot2)
setwd("~/Desktop/My-repo-master/My-repo/Graphs")
<<<<<<< HEAD
tiff("heatmap.tf.clustered.family.tiff", units="in", width=10, height=40, res=300)
# insert ggplot code

=======
tiff("heatmap TF with clustered family.tiff", units = "in", height = 40, width = 10, res = 500)
>>>>>>> ddb362d7cbac827fa2c3dc5c6d5b65b7bea9ce3d
ggplot(plotdata_melt, aes(x = Time, y = Protein, fill = value)) +
  geom_raster() +
  scale_fill_distiller(palette = "RdBu") +
  ggtitle("Heatmap of transcription factors")
<<<<<<< HEAD

#ggsave("TF.heatmap.arranged.clustered.pdf", height = 40, width = 10)

dev.off()
```

=======
dev.off()
#ggsave("TF.heatmap.arranged.clustered.png", height = 40, width = 10)
```


>>>>>>> ddb362d7cbac827fa2c3dc5c6d5b65b7bea9ce3d
```{r}
group1 = "100"
group2 = "110"
group3 = "101"
group4 = "010"
group5 = "001"
group6 = "1-10"
group7 = "10-1"
group8 = "-110"
group9 = "-100"
group10 = "-1-10"
group11 = "-1-1-1"
group12 = "-10-1"
<<<<<<< HEAD
group13 = "0-1-1"
group14 = "00-1"
group15 = "0-10"
=======
group13 = "0-10"
group14 = "0-1-1"
group15 = "00-1"
>>>>>>> ddb362d7cbac827fa2c3dc5c6d5b65b7bea9ce3d
```

```{r}
extract = function(data, term){
  group_tf = c()
  for(i in 1:length(data[[1]])){
<<<<<<< HEAD
    if(paste(as.character(data[i, 1:3]), collapse = "") == term){
=======
    if(paste(as.character(data[i,1:3]), collapse = "") == term){
>>>>>>> ddb362d7cbac827fa2c3dc5c6d5b65b7bea9ce3d
      group_tf = c(group_tf, row.names(data)[i])
    }
  }
  return(group_tf)
}
```

```{r}
group1_tf = extract(zincfinger, term = group1)
group2_tf = extract(zincfinger, term = group2)
group3_tf = extract(zincfinger, term = group3)
group4_tf = extract(zincfinger, term = group4)
group5_tf = extract(zincfinger, term = group5)
group6_tf = extract(zincfinger, term = group6)
group7_tf = extract(zincfinger, term = group7)
group8_tf = extract(zincfinger, term = group8)
group9_tf = extract(zincfinger, term = group9)
group10_tf = extract(zincfinger, term = group10)
group11_tf = extract(zincfinger, term = group11)
group12_tf = extract(zincfinger, term = group12)
group13_tf = extract(zincfinger, term = group13)
group14_tf = extract(zincfinger, term = group14)
group15_tf = extract(zincfinger, term = group15)
```

```{r}
new_order = c(group15_tf, group14_tf, group13_tf, group12_tf,
              group11_tf, group10_tf, group9_tf, group8_tf, group7_tf, group6_tf,
              group5_tf, group4_tf, group3_tf, group2_tf, group1_tf)
```

```{r}
zincfinger = zincfinger[new_order,]
<<<<<<< HEAD
=======
```


```{r}
group1 = "100"
group2 = "101"
group3 = "110"
group4 = "111"
group5 = "010"
group6 = "001"
group7 = "1-11"
group8 = "1-10"

```

```{r}
extract = function(data, term){
  group_tf = c()
  for(i in 1:length(data[[1]])){
    if(paste(as.character(data[i,1:3]), collapse = "") == term){
      group_tf = c(group_tf, row.names(data)[i])
    }
  }
  return(group_tf)
}
```

```{r}
group1_tf = extract(TF_bZIP, term = group1)
group2_tf = extract(TF_bZIP, term = group2)
group3_tf = extract(TF_bZIP, term = group3)
group4_tf = extract(TF_bZIP, term = group4)
group5_tf = extract(TF_bZIP, term = group5)
group6_tf = extract(TF_bZIP, term = group6)
group7_tf = extract(TF_bZIP, term = group7)
group8_tf = extract(TF_bZIP, term = group8)

```

```{r}
new_order = c(group8_tf, group7_tf, group6_tf,
              group5_tf, group4_tf, group3_tf, group2_tf, group1_tf)
```

```{r}
TF_bZIP = TF_bZIP[new_order,]
```

```{r}
group1 = "100"
group2 = "-100"
group3 = "0-10"
```

```{r}
group1_tf = extract(Tbox, term = group1)
group2_tf = extract(Tbox, term = group2)
group3_tf = extract(Tbox, term = group3)
```

```{r}
new_order = c(group3_tf, group2_tf, group1_tf)
```

```{r}
Tbox = Tbox[new_order,]
```

```{r}
group1 = "100"
group2 = "110"
group3 = "101"
group4 = "010"
group5 = "001"
group6 = "1-10"
group7 = "11-1"
group8 = "-110"
group9 = "-100"   
group10 = "-10-1"
group11 = "-1-10"
group12 = "0-1-1"
group13 = "0-10" 
group14 = "00-1"
group15 = "-1-1-1"
``` 

```{r}
group1_tf = extract(Homeobox, term = group1)
group2_tf = extract(Homeobox, term = group2)
group3_tf = extract(Homeobox, term = group3)
group4_tf = extract(Homeobox, term = group4)
group5_tf = extract(Homeobox, term = group5)
group6_tf = extract(Homeobox, term = group6)
group7_tf = extract(Homeobox, term = group7)
group8_tf = extract(Homeobox, term = group8)
group9_tf = extract(Homeobox, term = group9)
group10_tf = extract(Homeobox, term = group10)
group11_tf = extract(Homeobox, term = group11)
group12_tf = extract(Homeobox, term = group12)
group13_tf = extract(Homeobox, term = group13)
group14_tf = extract(Homeobox, term = group14)
group15_tf = extract(Homeobox, term = group15)
```

```{r}
new_order = c(group15_tf, group14_tf, group13_tf, group12_tf,
              group11_tf, group10_tf, group9_tf, group8_tf, group7_tf, group6_tf,
              group5_tf, group4_tf, group3_tf, group2_tf, group1_tf)
```

```{r}
Homeobox = Homeobox[new_order,]
```


```{r}
group1 = "100"
group2 = "110"
group3 = "010"
group4 = "001"
group5 = "1-10"
group6 = "-110"
group7 = "-100"

```


```{r}
group1_tf = extract(Fork_head, term = group1)
group2_tf = extract(Fork_head, term = group2)
group3_tf = extract(Fork_head, term = group3)
group4_tf = extract(Fork_head, term = group4)
group5_tf = extract(Fork_head, term = group5)
group6_tf = extract(Fork_head, term = group6)
group7_tf = extract(Fork_head, term = group7)

```


```{r}
new_order = c(group7_tf, group6_tf,
              group5_tf, group4_tf, group3_tf, group2_tf, group1_tf)
```

```{r}
Fork_head = Fork_head[new_order,]
```

```{r}
group1 = "100"
group2 = "110"
group3 = "010"
group4 = "101"
group5 = "001"
group6 = "0-10"
group7 = "-1-10"
group8 = "-10-1"
group9 = "00-1"   
``` 

```{r}
group1_tf = extract(bHLH, term = group1)
group2_tf = extract(bHLH, term = group2)
group3_tf = extract(bHLH, term = group3)
group4_tf = extract(bHLH, term = group4)
group5_tf = extract(bHLH, term = group5)
group6_tf = extract(bHLH, term = group6)
group7_tf = extract(bHLH, term = group7)
group8_tf = extract(bHLH, term = group8)
group9_tf = extract(bHLH, term = group9)
```

```{r}
new_order = c(group9_tf, group8_tf, group7_tf, group6_tf,
              group5_tf, group4_tf, group3_tf, group2_tf, group1_tf)
```

```{r}
bHLH = bHLH[new_order,]
>>>>>>> ddb362d7cbac827fa2c3dc5c6d5b65b7bea9ce3d
```
