Plot the heatmap of all the DE TFs from E11 to E14


```{r}
e12.e11.de.tf = read.csv("e12_e11_de_tf_0.5.csv", header = TRUE, stringsAsFactors = FALSE)
e12.e11.de.tf = e12.e11.de.tf[, c(3, 9)]
e13.e12.de.tf = read.csv("e13_e12_de_tf_0.5.csv", header = TRUE, stringsAsFactors = FALSE)
e13.e12.de.tf = e13.e12.de.tf[, c(3, 9)]
e14.e13.de.tf = read.csv("e14_e13_de_tf_0.5.csv", header = TRUE, stringsAsFactors = FALSE)
e14.e13.de.tf = e14.e13.de.tf[, c(3, 9)]
```

Merge data by the "Symbol" column. Missing values are replaced by "NA"
```{r}
e12.e11_e13.e12 = merge(e12.e11.de.tf, e13.e12.de.tf, by = "Symbol", all = TRUE)
merged_data = merge(e12.e11_e13.e12, e14.e13.de.tf, by = "Symbol", all = TRUE)
head(merged_data)
```
Change the column names

```{r}
colnames(merged_data) = c("Symbol", "E12/E11", "E13/E12", "E14/E13")
head(merged_data)
```

Function to convert positive logFC to 1 and negative logFC to -1. NA is assigned to 0

```{r}
up.down = function(logfc){
  if(is.na(logfc)){
    return(0)
  }else if(logfc < 0){
    return(-1)
  }else{return(1)}
}
```

Create a data frame suitable for plotting

```{r}
df.for.heatmap = data.frame(merged_data$Symbol)
df.for.heatmap$'E12/E11' = sapply(merged_data[['E12/E11']], up.down)
df.for.heatmap$'E13/E12' = sapply(merged_data[['E13/E12']], up.down)
df.for.heatmap$'E14/E13' = sapply(merged_data[['E14/E13']], up.down)
head(df.for.heatmap)
```
write this table to file
```{r}
write.csv(df.for.heatmap, "df_for_heatmap.csv")
```

Load necessary library

```{r}
library(ggplot2)
library(reshape2)
```
In order to use hclust to organize the order of gene, it is needed to make Symbol column
the row.names

```{r}
row.names(df.for.heatmap) = df.for.heatmap$merged_data.Symbol
df.for.heatmap = df.for.heatmap[, -1]
head(df.for.heatmap)
```

Calculate the order of gene using hclust
```{r}
row.order = hclust(dist(df.for.heatmap))$order
df.for.heatmap.new = df.for.heatmap[row.order,]
```

Now melt the data frame
```{r}
df.for.heatmap.molten = melt(as.matrix(df.for.heatmap.new))
names(df.for.heatmap.molten)[c(1,2)] = c("Protein", "Time")
```

Now plot the graph
```{r}
ggplot(df.for.heatmap.molten, aes(x = Time, y = Protein, fill = value)) +
  geom_raster() +
  scale_fill_distiller(palette = "RdBu") +
  ggtitle("Heatmap of transcription factors")

ggsave("TF.heatmap.not.arranged.png", height = 40, width = 10)
```
